---
title: "LDEO June-October EDA"
output: html_document
date: "2023-03-30"

---

```{r setup, include=FALSE, echo=FALSE, eval=TRUE}

require("knitr")
opts_knit$set(root.dir = "C:\\Users\\roseh\\Desktop\\NYBG_R\\Lamont\\data")
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


```{r echo=FALSE, eval=TRUE}
library(tidyverse)
library(lubridate)
library(GGally)
library(rstudioapi)

```



```{r echo=FALSE, eval=TRUE}
df <- read_csv("C:\\Users\\roseh\\Desktop\\NYBG_R\\Lamont\\data\\lamont_all_main.csv")
# convert data types
df$date <- mdy(df$date)
df$type <- as.factor(df$type)

df <- df %>%
  filter(lin_flux < 40)

# create df of mean values
df.m <- df %>%
  group_by(date,type) %>%
  summarize(mean_res = mean(lin_flux,na.rm=T),
            mean_temp = mean(chamber_temp,na.rm=T),
            mean_ndvi = mean(hh_ndvi,na.rm=T),
            mean_soil = mean(soil_temp,na.rm=T))

```


```{r echo=FALSE, eval=TRUE}
# function to plot temp v respiration
all_data_plot <- function(df,x,xlab){
  ggplot(df, aes(x=x,y=lin_flux)) +
    geom_point(aes(x=x,y=lin_flux,color=type)) + 
    geom_smooth() +
    ggtitle('Lamont All Data, June-Oct.') +
    xlab(xlab) +
    ylab('micromol CO2/m^2/s')
}
# function to plot mean temps v mean respiration
mean_data_plot <- function(df,x,xlab){
  ggplot(df, aes(x=x,y=mean_res)) +
    geom_point(aes(x=x,y=mean_res,color=type)) + 
    geom_smooth() +
    ggtitle('Lamont Mean Values, June-Oct.') +
    xlab(xlab)+
    ylab('micromol CO2/m^2/s')
}

# function to plot facet wrap
facet_plot <- function(df,x,xlab){
  ggplot(df, aes(x=x,y=lin_flux)) +
    geom_point(aes(x=x,y=lin_flux,color=type)) + 
    geom_smooth() +
    facet_wrap(~type,nrow=2) +
    ggtitle('Lamont All Data, June-Oct') +
    xlab(xlab)+
    ylab('micromol CO2/m^2/s')
}

facet_means <- function(df,x,xlab){
  ggplot(df, aes(x=x,y=mean_res)) +
    geom_point(aes(x=x,y=mean_res,color=type)) + 
    geom_smooth() +
    facet_wrap(~type,nrow=2) +
    ggtitle('Lamont Mean Values, June-Oct') +
    xlab(xlab)+
    ylab('micromol CO2/m^2/s')
}

```

```{r echo=FALSE, eval=TRUE}

# function to make boxplots of temperature and respiration by type
make_boxplot <- function(df,y,ylab){
  ggplot(df)+
    geom_boxplot(aes(x=type, y=y,fill=type)) +
    ggtitle('Lamont All Data, June-Oct ' , ylab) +
    xlab('Type')+
    ylab(ylab)
}

```



### Chamber Temperature v. CO2 

```{r echo=FALSE, eval=TRUE}

# plot chamber temp and soil temp v respiration
all_data_plot(df,df$chamber_temp,'Chamber Temperature')
mean_data_plot(df.m,df.m$mean_temp,'Chamber Temperature')


```

### Soil Temperature v. CO2 

```{r echo=FALSE, eval=TRUE}

all_data_plot(df,df$soil_temp,'Soil Temperature')
mean_data_plot(df.m,df.m$mean_soil,'Soil Temperature')


```

### Soil Temperature v. Respiration by Collar Type

```{r echo=FALSE, eval=TRUE}
# plot temp v. resp by type
facet_plot(df,df$soil_temp,'Soil Temperature')

```

### Chamber Temperature v. Respiration by Collar Type

```{r echo=FALSE, eval=TRUE}
facet_plot(df,df$chamber_temp,'Chamber Temperature')
```

### Chamber Temperature by Collar Type

```{r echo=FALSE, eval=TRUE}
df2 <- df %>%
  mutate(type=fct_reorder(.f = type, .x = chamber_temp, .fun = median, na.rm = T))
make_boxplot(df2,df2$chamber_temp,'Chamber Temperature')
```

### Soil Temperature by Collar Type

```{r echo=FALSE, eval=TRUE}

df4 <- df %>%
  mutate(type=fct_reorder(.f=type,.x=soil_temp,.fun=median,na.rm=T))
make_boxplot(df4,df4$soil_temp,'Soil Temperature')

```

### Respiration by Collar Type

```{r echo=FALSE, eval=TRUE}

df3 <- df %>%
  mutate(type=fct_reorder(.f = type, .x = lin_flux, .fun = median, na.rm = T))
make_boxplot(df3,df3$lin_flux,'Respiration')

```

### Scatterplot Matrix - correlations between all variables

```{r echo=FALSE, eval=TRUE}
m.res <- df.m$mean_res
m.temp <- df.m$mean_temp
m.ndvi <- df.m$mean_ndvi
m.soil <- df.m$mean_soil
df5 <- tibble(m.res,m.temp,m.ndvi,m.soil)
ggpairs(df5)
```

### NDVI v. Respiration

```{r echo=FALSE, eval=TRUE}

ndvi <- df %>%
  filter(!is.na(hh_ndvi))

ggplot(ndvi,aes(x=hh_ndvi,y=lin_flux)) +
  geom_point(aes(x=hh_ndvi,y=lin_flux,color=type)) +
  ylab('Respiration') +
  xlab('NDVI') +
  ggtitle('NDVI v. Respiration') 

ggplot(ndvi,aes(x=hh_ndvi,y=lin_flux)) +
  geom_point(aes(x=hh_ndvi,y=lin_flux,color=type)) +
  ylab('Respiration') +
  xlab('NDVI') +
  ggtitle('NDVI v. Respiration, by collar type') +
  facet_wrap(~type,nrow=2)


```



### Fitting multiple regression models

Model 1. Linear Model: respiration = b0 + chamber temperature + ndvi

```{r echo=FALSE, eval = TRUE}

mod1 <- lm(mean_res ~ mean_temp + mean_ndvi,df.m)
summary(mod1) 

```

Model 2. Linear model: respiration = b0 + soil temperature + ndvi

```{r}
mod2 <- lm(mean_res ~ mean_soil + mean_ndvi,df.m)
summary(mod2)
```

Model 3. Log transformation: respiration = b0 + log(soil temp) + log(ndvi)

```{r}

mod3 <- lm(mean_res ~ I(log(mean_soil)) + mean_ndvi,df.m)
summary(mod3) 
```

Model 4. Exponential model: respiration = b0 + exp(soil temperature + ndvi)

```{r}

mod4 <- lm(log(mean_res) ~ mean_soil + mean_ndvi,df.m) 
summary(mod4)

```

### Fitting simple regression models

Model 5. Quadratic model: respiration = b0 + soil temperature + (soil temperature)^2

```{r}

mod5 <- lm(mean_res ~ mean_soil + I(mean_soil**2),df.m) 
summary(mod5)
```
Model 6: Polynomial model: respiration = b0 + soil temp + soil temp^2 + soil temp^3

```{r}
library(splines)

mod6 <- lm(mean_res ~ ns(mean_soil,3),df.m)
summary(mod6)
```
```{r}
library(modelr)

grid <- df.m%>% 
  data_grid(mean_soil = mean_soil) %>% 
  gather_predictions(mod5, mod6, .pred = "mean_res")

ggplot(df.m, aes(mean_soil, mean_res)) + 
  geom_point() +
  geom_line(data = grid, colour = "red") +
  facet_wrap(~ model)
```
```{r}
mod7 <- lm(log(mean_res) ~ mean_soil + I(mean_soil**2) + mean_ndvi,df.m)
summary(mod7)
```






